# R packages

```{r}
knitr::opts_chunk$set(eval = FALSE)
```

>"R packages are collections of functions and data sets developed by the community. They increase the power of R by improving existing base R functionalities, or by adding new ones."^[https://www.datacamp.com/community/tutorials/r-packages-guide] 

Basically, R packages are nothing but collections of functions bundled together in a way that makes sense. Like different cookbooks that only contain recipes for a particular kind of food. They can be installed from many different sources which will be explored below. 

## From CRAN

[The Comprehensive R Archive Network](What are R and CRAN?) (CRAN) package repository features 18,000+ R packages. Here's the list of [Available CRAN Packages By Name](https://cran.r-project.org/web/packages/available_packages_by_name.html). Most general purpose packages can be found here, however due to reasons, some packages are only available from other sources. 

As a first example, we will download the [tidyverse](https://www.tidyverse.org/), a collection of R packages for data science. 

>"The tidyverse is an opinionated collection of R packages designed for data science. All packages share an underlying design philosophy, grammar, and data structures."

We can install packages from CRAN with the `install.packages()` function like this:
```{r install tidyR, eval=FALSE}
install.packages("tidyverse")
```

Note, that we can also download more than one package at once using a vector (`c()`) containing all package names:
```{r install more, eval=FALSE}
install.packages(c("BiocManager", "devtools"))
```

`BiocManager` as well as `devtools` will be used in the following to download R packages from other sources.


## From Bioconductor

The [Bioconductor](https://bioconductor.org/) is a collection of R packages for bioinformatics purposes. The first packages we will need from the Bioconductor will be downloaded with the `install()` function from the `BiocManager` package: 
```{r bioc, message=F}
BiocManager::install(c("fgsea",
                       "org.Hs.eg.db",
                       "UniProt.ws"))
```



## From GitHub and others sources

Another important source of R packages is GitHub. GitHub is not just the place where most R packages are being developed before they are put to repositories such as CRAN or Bioconductor, many other packages including the `pOmics3` packages can be installed from it as well. :)

Here is an example with the `install_github()` function from the `devtools` package which will be further described in [Setup pOmics3](#02_setup-pOmics3).
```{r install pOmics}
devtools::install_github("nicohuttmann/pOmics3")
```


There are some other sources for R packages, which we want to use. Here, we are downloading the `disgenet2r` package to query the [DisGeNET](https://www.disgenet.org/) database for gene disease associations:
```{r install disgenet2r}
devtools::install_bitbucket("ibi_group/disgenet2r")
```


