# (PART\*) Data Analysis Basics {.unnumbered}

# Project setup

```{r}
knitr::opts_chunk$set(eval = FALSE)
```


## Scripts


We should start with a place to save our code





## Data Structure

A good idea for every project is to have a simple structure to store your files. Here, I suggest having three main folders:

1.  Data to store all you raw data and other files that go into your analysis. In addition, you can create a sub folder to store all your RData files.
2.  Scripts to store all your analysis scripts and custom functions written for a specific analysis. 
3.  Output to store all files, plots and tables that come out of your analysis. Depending on the project size, sub folders for individual figures in a manuscript or file types could be useful.

```{r Create folders, eval=FALSE}
# ---- Setup folders ---- 
dir.create("Data")
dir.create("Data/raw")
dir.create("Data/RData")

dir.create("Scripts")

dir.create("Output")
```




Now you can add your raw data etc. In this example, we will work with this data frame: ... We can visualize the folder structure with the `fs::dir_tree` function.


```{r show folders and files, echo=FALSE}
fs::dir_tree(regexp = "Data|Scripts|Output")
```


