
# (PART\*) Data Analysis {.unnumbered}


# Data normalization 



## limma normalization methods {.tabset}




```{r}
data_norm <- data %>% 
  transpose_tibble() %>% 
  tibble2matrix() %>% 
  limma::normalizeBetweenArrays(method = "scale") %>% 
  matrix2tibble("variables") %>% 
  transpose_tibble()
```



## vsn normalization {.tabset}






## pqn normalization (implemented by pOmics3)




## Group-wise normalization {.tabset}


### base {-}

```{r}
data_norm <- data %>% 
  mutate(group = groups[observations], .after = "observations")
split(.$group) %>% 
  map(\(x) {
    x %>% select(-group) %>% 
      transpose_tibble() %>% 
      tibble2matrix() %>% 
      limma::normalizeBetweenArrays(method = "scale") %>% 
      matrix2tibble("variables") %>% 
      transpose_tibble()
  }) %>% 
  bind_rows() %>% 
  arrange(match(observations, names(groups)))
```


## Principal component analysis (PCA)

(How to use PCA)



### base {-}

```{r}
list_pca <- list()

list_pca[["prcomp"]] <- 
  get_data_frame(which = "median", 
                 dataset = "protein") %>% 
  mutate(across(where(is.numeric), \(x) c(scale(x)))) %>% 
  tibble2data_frame() %>% 
  prcomp()
```

```{r}
list_pca[["p12"]] <- list_pca[["prcomp"]][["x"]] %>% 
  matrix2tibble(to.row.names = "observations") %>%  
  mutate(group = groups[observations], .after = "observations") %>% 
  ggplot(aes(x = PC1, y = PC2, color = group)) +
  geom_point() +
  theme_classic() +
  scale_color_manual(values = Info$freezing_colors) +
  coord_fixed() + 
  xlab("PC1") + 
  ylab("PC2")
```


```{r}
list_pca[["p12"]] +
  xlab(paste0(
    "PC1 (", 
    round(
      with(
        list(x = list_pca[["prcomp"]][["sdev"]]), 
        x^2 / sum(x^2))[1] * 100, 1), "%)")) + 
  ylab(
    paste0(
      "PC2 (", 
      round(
        with(
          list(x = list_pca[["prcomp"]][["sdev"]]), 
          x^2 / sum(x^2))[2] * 100, 1), "%)"))
```


### pOmics3 {-}

```{r}
Analysis[["PCA"]] <- list()

Analysis[["PCA"]][["prcomp"]] <- 
  get_data_frame(which = "median", 
                 variables = , 
                 observations = , 
                 dataset = "protein") %>% 
  mutate(across(where(is.numeric), \(x) c(scale(x)))) %>% 
  tibble2data_frame() %>% 
  prcomp()
```



```{r}
Analysis[["PCA"]][["p12"]] <- 
  (Analysis[["PCA"]][["prcomp"]][["x"]] %>% 
     matrix2tibble(to.row.names = "observations") %>%  
     add_observations_data("group", dataset = "protein") %>% 
     add_observations_data("group2", dataset = "protein") %>% 
     ggplot(aes(x = PC1, y = PC2, color = group, shape = group2)) +
     geom_point() +
     theme_classic() +
     theme(panel.background = element_rect(color = "black")) +
     #scale_color_manual(values = Info$color_scheme) +
     coord_fixed()) %>% 
  .set_PCA_labs() %>% 
  .set_continuous_axes(axis.unit.ratio = 1)
```


### Combined PCA plots


```{r}
wrap_plots(list_pca[["p12"]], Analysis[["PCA"]][["p34"]], guides = "collect")
```


```{r}
wrap_plots(Analysis[["PCA"]][["p12"]], Analysis[["PCA"]][["p34"]], guides = "collect")
```


### How to get creative

PCA can be used on subsets of samples or on a subset of proteins of interest.




## Differential abundace analysis with limma


```{r}
# limma
list_limma <- list()

# Make eset from tibble
list_limma[["eset"]] <- 
  get_data_frame("median",
                 variables = , 
                 observations = ,
                 dataset = "protein") %>% 
  mutate(across(where(is.numeric), log2)) %>% 
  transpose_tibble() %>% 
  tibble2matrix() %>% 
  Biobase::ExpressionSet()
```


```{r}
# Describe experimental groups
list_limma[["design"]] <- model.matrix(
  ~0+factor(
    get_observations_data(
      which = "group", 
      observations = ,
      output.type = "vector", 
      dataset = "protein"), 
    levels = c("A", "B", "C"))) # order of groups

colnames(list_limma[["design"]]) <- c("A", "B", "C") # same as levels above

# Do first linear model fit
list_limma[["fit"]] <- lmFit(list_limma[["eset"]], list_limma[["design"]])

# Describe comparisons (contrasts)
list_limma[["contrast.matrix"]] <- 
  makeContrasts(B-A, C-A, B-C, # describe desired comparison
                levels = list_limma[["design"]])

# Compute contrasts
list_limma[["fit2"]] <- contrasts.fit(list_limma[["fit"]], list_limma[["contrast.matrix"]])

# Apply empirical bayes statistics
list_limma[["fit2_eBayes"]] <- eBayes(list_limma[["fit2"]])
```





```{r}
p.threshold <- 0.05
fc.threshold.hit <- 1

p.adjust.method <- "BH"

# Tidy up results
list_limma[["results"]] <- biobroom::tidy.MArrayLM(list_limma[["fit2_eBayes"]]) %>% 
  # do p-value correction
  group_by(term) %>% 
  mutate(p.adjust = p.adjust(p.value, method = p.adjust.method), .after = "p.value") %>% 
  ungroup()
# 
mutate(regulation = "none",
       regulation = ifelse(p.adjust < p.threshold & estimate >= fc.threshold.hit,
                           "hit.up",
                           regulation),
       regulation = ifelse(p.adjust < p.threshold & estimate <= - fc.threshold.hit,
                           "hit.down",
                           regulation)) %>% 
  split(.$term)
```

Add candidates if wanted
```{r}

         # regulation = ifelse(p.value < p.threshold & estimate >= fc.threshold.candidate,
         #                     "candidate.up",
         #                     regulation),
         # regulation = ifelse(p.value < p.threshold & estimate <= - fc.threshold.candidate,
         #                     "candidate.down",
         #                     regulation),
```

Add protein and gene names.

```{r}
list_limma[["results"]] <- list_limma[["results"]] %>% 
  rename(variables = gene) %>% 
  mutate(Protein.name = pull(Info[["Databases"]][["UniProt"]], 
                             "Protein names", 
                             "Entry")[variables], 
         Gene.name = pull(Info[["Databases"]][["UniProt"]], 
                          "Gene Names", 
                          "Entry")[variables], 
         .after = 1) %>% 
```





```{r}
list_limma[["p"]] <- 
  names(list_limma[["results"]]) %>% 
  map(\(x) 
      ggplot(list_limma[["results"]][[x]], 
             aes(x = estimate, 
                 y = -log10(p.value), 
                 color = regulation, 
                 Protein.name = Protein.name, 
                 Gene.name = Gene.name)) +
        geom_point() +
        scale_color_manual(values = c(none = "grey", 
                                      hit.up = "red", 
                                      hit.down = "blue")) +
        ggtitle(x))

Analysis[["Liver_total_limma"]] <- list_limma
```


```{r}
patchwork::wrap_plots(Analysis[["Liver_total_limma"]][["p"]], guides = "collect")
```



```{r}
plotly::ggplotly(Analysis[["Liver_total_limma"]][["p"]][[1]])
```











